### 2022-11-08 presentation/robustMR.qmd
revealjs presentation 2022-11-08

---
title: "Assessment of simulation procedures and robust methods for Mendelian randomization analysis"
author: "Christoph Reich"
date: "`r Sys.Date()`"
title-slide-attributes: 
  data-background-image: ../templates/bg1.png
  data-background-size: cover  # default
  data-background-opacity: "1"  # from 0-1
format: 
  revealjs:
    slide-number: true
    scrollable: true
    navigation-mode: vertical
    chalkboard:   # kreidetafel
      theme: whiteboard
      boardmarker-width: 5
      buttons: false
    incremental: false  # display bullet lists all at once
    preview-links: auto
    logo: ../templates/mederlab_logo.png
    css: ../templates/main.css
    footer: "robustMRmethods"
    theme: [simple, ../templates/custom.scss]
    #self-contained: true  # for distribution (slows rendering), not compatible with chulkboarddic
# reference-doc: templates/20210129_mederslidemaster.pptx  # only for pptx rendering
bibliography: ../references/references.bib
---

```{r}
#| label: dependencies
#| include: false
#| warning: false

rm(list=ls())

knitr::opts_chunk$set(echo = FALSE)

library(dplyr)
library(tidyr)
library(stringr)
library(tibble)
library(ggplot2)
library(scales)
library(ggthemes)
#library(biomaRt)  # different script
library(kableExtra)
library(MendelianRandomization)
library(mr.raps)
library(MRMix)
library(MRPRESSO)
library(penalized)
library(conflicted)
conflict_prefer("select", "dplyr")
conflict_prefer("filter", "dplyr")

theme_set(ggthemes::theme_few()) 
```

```{r}
#| label: load-data
#| include: false
#| warning: false

my_data <- as_tibble(read.delim("../data-boekstegers_felix/GST_BMI_association_results.txt"))
# GBC: new data 2022-11-04
GBC_dat <- as_tibble(read.delim("../data-boekstegers_felix/C23C24_BMI_association_results.txt"))

GBC_dat <- GBC_dat %>% 
  rename(eaf_gbc=eaf_out,
         beta_gbc=beta_out,
         se_gbc=se_out,
         p_value_gbc=p_value_out) %>% 
  select(-c(chr, pos_hg19, pos_hg38, a1, a2, exposure, outcome))
my_data <- my_data %>% left_join(GBC_dat, by =c("SNP", "dummyID"))
```

## Get gene names from rs-ids

```{r}
#| label: FTO-annotation
#| tbl-cap: "SNP associated with FTO gene"
#| tbl-colwidths: [60,40]

# see script biomart_search.R
rsIds_GeneNames <- as_tibble(readRDS("../data/biomartBMIsearch.rds"))

filter_FTO <- rsIds_GeneNames %>% 
  filter(str_detect(associated_gene, pattern = "FTO")) %>% 
  distinct(refsnp_id, .keep_all = TRUE)

t(filter_FTO) %>% kable() %>% 
  kable_styling(font_size=18)
```

:::: {.columns}

::: {.column width="50%"}
* from the 284 SNPs only one was associated with the FTO gene
:::

::: {.column width="50%"}
![](../img/ensembl_logo.png)

:::

::::

## Simulating on empirical data

### Calculate approximate SEs for decreasing sample sizes and corresponding p-vals

* se, pval and study size depend on each other, within each study
* selection of IVs based on exposure study (BMI in our case)
* Summary statistics on genetic associations with BMI based on results of the GIANT + GERA studies [@Hoffmann_2018] with $N=100,418 + 234,069 =334487$  and $SNPs=289$

## Steps

1. Calculate more accurates SEs using the betas and Pvals
2. Calculate the approximate SEs for decreasing samples sizes
3. Calculate the probability values considering the estimated betas and the corresponding SEs
4. Exclude the IV if Pval > genome-wide significance threshold  (5 × 10−8) **NEW 5x10^(-6))**
5. Calculate cumulative explained variances
6. Represent figure similar to Fig1 in Chatterjee’s paper @QiChatterjee_2021

```{r}
#| label: z-statistic

### justo approach email 30-10-2022  ---------------------------------------
## LOOP simulation ----------------------------------------------
n.orig <- 334487
n <- c(1000, 2000, 5000, seq(10000, 330000, by=10000), 334487)

# create vectors to store pvals_n in
colnames_pval <- c()
for (i in 1:length(n)) {
  colnames_pval[i] <- paste0("p_value_exp_sample_", format(n[i], scientific = FALSE))
}
## initialize
se_update <- c()
pvals_df <- data.frame(matrix(ncol = length(colnames_pval), nrow =  nrow(my_data)))
colnames(pvals_df) <- colnames_pval
for (i in 1:nrow(my_data)) {
  # 1) Calculate more accurates SEs using the provided betas and Pvals 
  se_update[i] <- my_data[["beta_exp"]][i]/-qnorm(p=(my_data[["p_value_exp"]][i]/2),mean=0)

  # 2) Calculate the approximate SEs for decreasing samples sizes
  ### approximation of sd (single value)
  sd <- se_update[i] * sqrt(n.orig)
  #### approximation of standard error simulating different sample size
  se_n <- sd/(sqrt(n))

  # 3) Calculate the probability values considering the estimated betas and the corresponding SEs
  for (j in 1:length(n)) {
    pvals_n <- 2*(1-pnorm(abs(my_data[["beta_exp"]][i])/abs(se_n[j]), mean=0, 
                          lower.tail = TRUE))
    pvals_df[i, j] <- pvals_n
  }
}

# 4. Exclude the IV if Pval > genome-wide significance threshold  (5 × 10−8)
my_data <- as_tibble(cbind(my_data, pvals_df))

pvals_df %>% 
  summarize(across(.cols = everything(), function(x) sum(x < 5*10^(-6)))) %>% 
  t() %>% 
  kable(caption = "Count of significant SNPs depending on sample size") %>% 
  kable_styling(font_size=18)
```

## Explained variance

* The variance in BMI explained by the IVs is calculated with the following formula: $explained\ variance = 2f(1-f)\beta^2$, where $f$ denotes the allele frequency and $\beta$ is the additive genetic effect. 
* The explained variance in liability to GSD and GBC is calculated as described by @So_2011.

```{r}
#| label: calculate-explained-variance

explained_variance_numeric <- function(eaf, beta) {
  # eaf = allele frequency
  # beta = additive genetic effect
  explained_variance_numeric <- 2*eaf *(1-eaf) * beta^2
  return(explained_variance_numeric)
}

explained_variance_binary <- function(PA,RR1,RR2,K) {
  # from So et al 2011
  # Calculates the variance in liability explained by a single biallelic loci.
  # PA: allele frequency of the risk allele (denoted A)
  # RR1: relative risk of Aa (one risk allele) compared to aa (no risk allele)
  # RR2: relative risk of AA (two risk alleles) compared to aa (no risk allele)
  # K:  overall probability of disease in population
  # Returns the variance explained (Vg) and the mean liability for each genotype (the overall liability is normalized to mean 0 and variance 1)
  Paa = (1-PA)^2
  PAa = 2*PA*(1-PA)
  PAA = PA^2
  muaa=0
  faa= K/(Paa + PAa*RR1 + PAA*RR2)
  fAa= RR1*faa
  fAA= RR2*faa 
  T = qnorm(1-faa) 
  muAa = T-qnorm(1-fAa)
  muAA = T-qnorm(1-fAA)
  mean.all= PAa*muAa+ PAA*muAA
  Vg= Paa*(muaa-mean.all)^2 + PAa*(muAa-mean.all)^2+ PAA*(muAA-mean.all)^2
  actual.Vg =  Vg/(1+Vg) 
  VR = 1-actual.Vg 
  actual.T = Paa*sqrt(VR)*qnorm(1-faa) + PAa*sqrt(VR)*qnorm(1-fAa) + PAA*sqrt(VR)*qnorm(1-fAA)
  actual.muaa = actual.T - sqrt(VR) * qnorm(1-faa)
  actual.muAa = actual.T - sqrt(VR) * qnorm(1-fAa)
  actual.muAA = actual.T - sqrt(VR) * qnorm(1-fAA)

  res <- list(Vg=actual.Vg,muaa=actual.muaa, muAa = actual.muAa, muAA=actual.muAA)
  res
}

## calculate explained variance
my_data <- my_data %>% 
  mutate(explained_variance_BMI = explained_variance_numeric(eaf = eaf_exp, beta = beta_exp), 
         explained_variance_GSD = explained_variance_binary(PA = eaf_out, 
                                                            RR1 = exp(beta_out), 
                                                            RR2 = exp(beta_out)^2, 
                                                            K = 0.1  # assumed prevalence in population
                                                            )$Vg,
         explained_variance_GBC = explained_variance_binary(PA = eaf_gbc, 
                                                            RR1 = exp(beta_gbc), 
                                                            RR2 = exp(beta_gbc)^2, 
                                                            K = 1.2/100000  # assumed prevalence in population
                                                            )$Vg
         )

## paper backing prevalence GSD
# https://www.ncbi.nlm.nih.gov/pmc/articles/PMC3343155/#:~:text=Gallstones%20are%20common.,develop%20stones%20at%20some%20time.&text=The%20majority%20will%20not%20develop,cholecystitis%2C%20cholangitis%2C%20or%20pancreatitis
## paper backing prevalence GSD (ASR/100.000 = 1.2 worldwide)
# https://www.wcrf.org/cancer-trends/gallbladder-cancer-statistics/

#explained_variance_binary(PA=0.3,RR1=1.3,RR2=1.3^2,K=0.01)
```

## Plot explained variance in BMI vs -log10 P-value from Hoffmann et al.

```{r}
#| label: explainedBMIvarianceVSpval
#|

my_data$gene_label <- NA
my_data$gene_label[my_data$SNP %in% filter_FTO$refsnp_id] <- "FTO"

# color labeling
my_data$fto_col_annot <- "No FTO variant"
my_data$fto_col_annot[my_data$SNP %in% filter_FTO$refsnp_id] <- "FTO variant"

ggplot(data = my_data, 
       mapping = aes(x=explained_variance_BMI*100, y= -log10(p_value_exp),  
                    col=fto_col_annot, label=gene_label)
       )+
  geom_point(alpha=0.6)+
  ggrepel::geom_text_repel(show.legend = FALSE)+
  xlab("Explained variance in BMI in %")+
  ylab("-log10 P-val")+
  labs(title = "Explained variance exposure",
       col='SNP'
       ) +
  theme_few()+
  scale_colour_few() -> fig00
fig00

ggsave(filename = "../output/plots/figure00.png", plot = fig00, 
       width = 10, height = 6, 
       units = "in"  # default
       )
```

## Plot Figure 1 from Qi&Chatterjee 2021

Here we show a plot similar to Figure 1 in [@QiChatterjee_2021]:

```{r get-data-for-fig1}
# 4.       Exclude the IV if Pval > genome-wide significance threshold  (5 × 10−8) **NEW 5x10^(-6))**
# 5.       Calculate cumulative explained variances
# initialize df
df_cum_expl_var <- tibble(n=n, 
                          cum_expl_varX = rep(NA, length(n)), 
                          cum_expl_varU = rep(NA, length(n)), 
                          cum_expl_varY = rep(NA, length(n)),
                          count_IVs = rep(NA, length(n))
)

for (i in 1:length(colnames_pval) ) {
  threshold <- 5*10^(-6)
  select_col <- colnames_pval[i]
  # calculate cum explained variance and count of IVs
  summarydat <- my_data %>% 
    filter(!!sym(select_col) < threshold) %>% 
    select(explained_variance_BMI, explained_variance_GSD, explained_variance_GBC) %>% 
    summarise(expl_variances =across(.cols = everything(), function(x) sum(x)),
              count_IVs = n())
  # fill up df
  df_cum_expl_var[i, 2] <- summarydat$expl_variances[1]
  df_cum_expl_var[i, 3] <- summarydat$expl_variances[2]
  df_cum_expl_var[i, 4] <- summarydat$expl_variances[3]
  df_cum_expl_var[i, 5] <- summarydat$count_IVs
}
```

```{r}
#| label: figure1QiChatterjee
# 6.       Represent figure similar to Fig1 in Chatterjee’s paper

# Dual Y axis: we have 284 IVs, scale them to 5% (y-axis [0;5%]): 3/284 
## https://r-graph-gallery.com/line-chart-dual-Y-axis-ggplot2.html
scale_factor <- 4.5/284
df_cum_expl_var <- df_cum_expl_var %>% 
  mutate(
    cum_expl_varX = cum_expl_varX *100, # in percent 
    cum_expl_varU = cum_expl_varU *100,
    cum_expl_varY = cum_expl_varY *100,
    count_IVs_scaled = count_IVs*scale_factor)

plot_data <- df_cum_expl_var %>% 
  pivot_longer(cols = c(cum_expl_varX, cum_expl_varU, count_IVs_scaled)) %>% 
  mutate(name = factor(name, labels = c("Number of IVs", "% var(U)", "% var(X)")))

ggplot(plot_data, aes(x=log10(n), y=value, col=name))+
  geom_point(alpha=0.6)+
  geom_line(alpha=0.6)+
  xlab("log10(N)")+
  #ylab("% variance explained by IVs")+
  labs(title = "Figure 1",
       col='', 
       caption = "Relationship between sample size, average number of instrumental variables (IVs) and variance of traits explained by the IVs"
       ) +
  scale_y_continuous(
    # Features of the first axis
    name = "% variance explained by IVs",
    # Add a second axis and specify its features
    sec.axis = sec_axis(~./scale_factor, name="Number of IVs")
  ) + 
  theme_few()+
  scale_colour_few() ->fig01

ggsave(filename = "../output/plots/figure01.png", plot = fig01, 
       width = 10, height = 6, 
       units = "in"  # default
       )
```

```{r updated-fig01}
plot_data1 <- df_cum_expl_var %>% 
  pivot_longer(cols = c(cum_expl_varX, cum_expl_varU, cum_expl_varY)) %>% 
  mutate(name = factor(name, labels = c("% var(U)", "% var(X)", "% var(Y)")))

N_annotation <- c(10000, 100000, 200000, 300000)
## plot
ggplot(plot_data1, aes(x=n, y=value, col=name))+
  geom_point(alpha=0.6)+
  geom_line(alpha=0.6)+
  geom_line(aes(x=n, y=count_IVs_scaled, col="Number of IVs"), linetype="dotdash")+
  scale_x_continuous(name = "Sample Size",
                     labels = comma, 
                     breaks = N_annotation, #c( 10000, n[7:length(n)]),
                     limits = c(10000, 350000))+
  #ylab("% variance explained by IVs")+
  labs(title = "Figure 1",
       col='', 
       caption = "Relationship between sample size, average number of instrumental variables (IVs) and variance of traits explained by the IVs"
       ) +
  scale_y_continuous(
    # Features of the first axis
    name = "% variance explained by IVs",
    # Add a second axis and specify its features
    sec.axis = sec_axis(~./scale_factor, name="Number of IVs")
  ) + 
  theme_few()+
  #scale_colour_few() +
  #scale_color_manual(name = "", values = c("% var(U)" = "red", 
  #                                         "% var(X)" = "blue", 
  #                                         "% var(Y)" = "darkblue", 
  #                                         "Number of IVs" = "black"
  #                                         ))
  #
  scale_color_manual(name = "", values = c("% var(U)" = ggthemes_data$few$colors$Dark[2,2][[1]], 
                                           "% var(X)" = ggthemes_data$few$colors$Dark[3,2][[1]], 
                                           "% var(Y)" = ggthemes_data$few$colors$Dark[4,2][[1]], 
                                           "Number of IVs" = "black"
                                           )) -> fig01.update
fig01.update
# scale_color_manual(name = "Y series", values = c("Y1" = "darkblue", "Y2" = "red"))
ggsave(filename = "../output/plots/figure01update.png", plot = fig01.update, 
       width = 10, height = 6, 
       units = "in"  # default
       )
```

## Different genetic architectures of the exposure

```{r MAF-Pval}
ggplot(data = my_data, aes(x=eaf_exp, y=-log10(p_value_exp)))+
  geom_point(alpha=0.6)+
  geom_hline(yintercept=-log10(median(my_data$p_value_exp)), col="red", alpha=0.6)+
  geom_vline(xintercept=median(my_data$eaf_exp), col="red", alpha=0.6)+
  xlab("Effect allele frequency for exposure")+
  ylab("-log10(P-value) ")+
  labs(title = "Figure 2",
       subtitle = "Define different genetic architectures",
       #col='', 
       #caption = "Relationship between sample size, average number of instrumental variables (IVs) and variance of traits explained by the IVs"
       ) +
  theme_few()+
  scale_colour_few() ->fig02
#fig02

ggsave(filename = "../output/plots/figure02.png", plot = fig02)
```

### MAF vs betas

```{r mafVSbetas}
plot_data_2 <- my_data %>% 
  mutate(maf_exp = ifelse(eaf_exp >0.5, 1-eaf_exp, eaf_exp)) %>% 
  select(maf_exp, beta_exp)

ggplot(data = plot_data_2, aes(x=maf_exp, y=abs(beta_exp)))+
  geom_point(alpha=0.6)+
  geom_hline(yintercept=median(abs(plot_data_2$beta_exp)), col="red", alpha=0.6)+
  geom_vline(xintercept=median(plot_data_2$maf_exp), col="red", alpha=0.6)+
  xlab("Minor allele frequency for exposure")+
  ylab("abs(beta_BMI)")+
  labs(title = "Figure 2-update",
       subtitle = "Define different genetic architectures",
       #col='', 
       #caption = "Relationship between sample size, average number of instrumental variables (IVs) and variance of traits explained by the IVs"
       ) +
  theme_few()+
  scale_colour_few() ->fig02.update
fig02.update
ggsave(filename = "../output/plots/figure02update.png", plot = fig02.update)
```

## Simulation Framework

* Estimate parameters for the simulation (pi1, varx, pi2, varu, muy, vary, pi3, pi4, delta0) based on real data
* Start with $N=10.000$ and $N=300.000$

![](../img/m4qi2021.png)

##

**References**

::: {#refs}
:::

**Software**

This analysis was carried out using the statistical software `r base::version$version.string` along with [RStudio](https://www.rstudio.com).
Apart from R's base functionality, the following packages were used: <br/>

```{r software, echo=FALSE}
# or faster with function...
installed.packages()[names(sessionInfo()$otherPkgs), "Version"]
```

##  {background-image="../templates/bg4.png"}

::: footer

:::