### 2022-12-16 scripts/test/MR-methods.Rmd
issue standard deviation for simulating from normal distribution detected

---
title: "MR-methods"
author: "Christoph Reich"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

---
# Perform MR-analysis

We first perform MR analysis.

```{r MendelianRandomization}
# create object
mr_object <- MendelianRandomization:: mr_input(
  # exposure
  bx = my_data$beta_exp, bxse = my_data$se_exp, 
  # outcome
  by = my_data$beta_out, byse = my_data$se_out, 
  exposure = "Body mass index",
  outcome = "Gallstone disease")

#perform MR-Robust
mrrobust <- mr_ivw(mr_object, "random", robust = TRUE)  # {MendelianRandomization}
#mrrobust

#perform MR-Egger
mregger <- mr_egger(mr_object)  # {MendelianRandomization}
#mregger

#perform weighted median
weighted_median <- mr_median(mr_object,  # {MendelianRandomization}
                             weighting = "weighted", 
                             iterations = 10000)
#weighted_median

#perform Mode based estimation
mode_based <- mr_mbe(mr_object,  # {MendelianRandomization}
       weighting = "weighted", 
       stderror = "delta", 
       phi = 1,
       seed = 19940407, 
       iterations = 10000, 
       distribution = "normal",
       alpha = 0.05)
#mode_based
```

```{r MRMix}
#perform MR-Mix  (Slob & Burgess squared their standard deviations`???); in vignette before analysis summary stats are standardized with `standardize` function
mrmix_est <- MRMix::MRMix(betahat_x = my_data$beta_exp, 
                          betahat_y = my_data$beta_out, 
                          sx = my_data$se_exp, 
                          sy = my_data$se_out)
#mrmix_est

mrmix_se <- MRMix::MRMix_se(betahat_x = my_data$beta_exp, 
                          betahat_y = my_data$beta_out, 
                          sx = my_data$se_exp, 
                          sy = my_data$se_out, 
                          theta = mrmix_est$theta, # estimate of causal effect, assuming the summary stats are standardized, theta represents increase in mean value of Y in s.d. unit of Y (for cont outcomes) or log-OR of Y (for binary outcomes)
                          pi0 = mrmix_est$pi0, 
                          sigma2 = mrmix_est$sigma2
                          )
#mrmix_se
```

```{r MRPRESSO}
#perform MR-PRESSO
# presso <- MRPRESSO:: mr_presso(BetaOutcome = "beta_out", 
#                     BetaExposure = "beta_exp", 
#                     SdOutcome = "se_out", 
#                     SdExposure = "se_exp", 
#                     OUTLIERtest = TRUE, 
#                     DISTORTIONtest = TRUE, 
#                     data = my_data, 
#                     NbDistribution = 1000,   # must be changed
#                     SignifThreshold = 0.05)
```

```{r MR-lasso}
#define function for MR-Lasso with heterogeneity criterion
MR_lasso<-function(betaYG, betaXG,sebetaYG){
  betaYGw = betaYG/sebetaYG # dividing the association estimates by sebetaYG is equivalent
  betaXGw = betaXG/sebetaYG # to weighting by sebetaYG^-2
  pleio = diag(rep(1, length(betaXG)))
  l1grid = c(seq(from=0.1, to=5, by=0.1), seq(from=5.2, to=10, by
  =0.2))

  # values of lambda for grid search
  l1grid_rse = NULL; l1grid_length = NULL; l1grid_beta = NULL; l1grid_se = NULL
  for (i in 1:length(l1grid)) {
    l1grid_which = which(attributes(penalized(betaYGw, pleio,
                                              betaXGw, lambda1=l1grid[i], trace=FALSE))$penalized==0)

    l1grid_rse[i] = summary(lm(betaYG[l1grid_which]~betaXG[l1grid_which]-1, 
                               weights=sebetaYG[l1grid_which]^-2))$sigma 
    l1grid_length[i] = length(l1grid_which)
    l1grid_beta[i] = lm(betaYG[l1grid_which]~betaXG[l1grid_which]-1,
                        weights=sebetaYG[l1grid_which]^-2)$coef[1]
    l1grid_se[i] = summary(lm(betaYG[l1grid_which]~betaXG[l1grid_which]-1, 
                              weights=
                                sebetaYG[l1grid_which]^-2))$coef[1,2]/
      min(summary(lm(betaYG[l1grid_which]~betaXG[l1grid_which]-1,
                     weights=sebetaYG[l1grid_which]^-2))$sigma, 1)
  }

  l1which_hetero = c(which(l1grid_rse[1:(length(l1grid)-1)]>1 & 
                             diff(l1grid_rse)>qchisq(0.95, df=1)/l1grid_length[2:length(l1grid)]), 
                     length(l1grid))[1]
  # heterogeneity criterion for choosing lambda
  l1hetero_beta = l1grid_beta[l1which_hetero] 
  l1hetero_se = l1grid_se[l1which_hetero]
  list(ThetaEstimate=l1hetero_beta, ThetaSE=l1hetero_se )
}

#perform MR-Lasso
mrlasso <- MR_lasso(betaYG = my_data$beta_out, betaXG = my_data$beta_exp, sebetaYG = my_data$se_out)
#mrlasso
```

```{r contamination-mixture}
contaminationmixture<-function(by, bx, byse){
  iters = 2001; theta = seq(from=-3, to=3, by=2/(iters-1))
  # if the causal estimate (and confidence interval) is not expected
  #     to lie between -1 and 1 then change from and to (and maybe
  #     increase iters)
  ratio = by/bx; ratio.se = abs(byse/bx); psi = 1.5*sd(ratio) 
  lik=NULL

  for (j1 in 1:iters) {
    lik.inc = exp(-(theta[j1]-ratio)^2/2/ratio.se^2) /sqrt(2*pi* ratio.se^2)
    lik.exc = exp(-ratio^2/2/(psi^2+ratio.se^2)) /(sqrt(2*pi*(psi^2+ ratio.se^2)))
    valid = (lik.inc>lik.exc)*1
    lik[j1] = prod(c(lik.inc[valid==1], lik.exc[valid==0])) 
    if (which.max(lik)==length(lik)) { valid.best = valid }
  }

  phi = ifelse(test = sum(valid.best)<1.5, 
               yes = 1, 
               no = 
                 max(
                  sqrt(sum(
                    ((ratio[valid.best==1]-weighted.mean(ratio[valid.best==1], ratio.se[valid.best==1]^-2))^2 * ratio.se[valid.best==1]^-2)) / 
                      (sum(valid.best) -1)),
                  1)
               )
  loglik = log(lik)
  whichin = which(2*loglik>(2*max(loglik)-qchisq(0.95, df=1)*phi^2))

  theta[which.max(loglik)] # estimate
  theta[whichin[1]] # lower limit of CI 
  theta[whichin[length(whichin)]] # upper limit of CI
  # return
  list(ThetaEstimate=theta[which.max(loglik)], ThetaLower=theta[ whichin[1]] , ThetaUpper= theta[whichin[length(whichin)]] )
}

#perform contamination mixture, note Slob & Burgess removed the 27th variable 
#   due to having a ratio to close to infty in their analysis
contam.mix <- contaminationmixture(by = my_data$beta_out, 
                                   bx = my_data$beta_exp, 
                                   byse = my_data$se_out)
#contam.mix
```

---